Dynamics of Colombo's Top: Generating Exoplanet Obliquties from Planet-Disc Interactions

### 2020-04-13 writeups/Su_disk_disp.tex
check in draft 2 maybe?

\begin{figure}
    \includegraphics[width=0.5\textwidth]{plots_diskdisp/2_cs_locs.png}
    \caption{Cassini state obliquities as a function of $\eta$ for $I =
    5^\circ$. The vertical dashed line indicates $\eta_{\rm c}$ ($= 0.766$ for
    $I = 5^\circ$), where CS1 and CS4 merge and annihilate, and the horizontal
    dashed lines indicate $\theta = I$ and $I - 180^\circ$, the asymptotic
    values for CSs 2 and 3 for $\eta \gg \eta_{\rm c}$.}\label{fig:cs_locs}
\end{figure}

@@

As shown in Appendix~\ref{ss:eigens}, libration about CS2 is slower than that
about CS1 or CS3. As such, it has the smallest characteristic frequency in the
system. The weak adiabaticity criterion can be recast as requiring that $\eta$
vary more slowly than this libration frequency \citep[see][]{ward2004II}, i.e.\
$t_{\rm d}^{-1} \lesssim \omega_{\rm lib} / 2\pi$. In terms of the dimensionless
parameter $\epsilon$
[Eq.~\eqref{eq:eps_def}], this adiabaticity condition is
\begin{equation}
    \epsilon \lesssim \frac{1}{2\pi}\sqrt{\eta\sin I \sin \theta_2
            \p{1 + \eta \sin I \csc^3 \theta_2}},
            \label{eq:ad_constr}
\end{equation}
where $\theta_2$ is the obliquity at CS2. This formula differs from that given
in \citet{millholland_disk}, where the $\csc^3\!\theta_2$ term is neglected and
the square root is missing\footnote{The missing $\csc^3 \theta_2$ term can be
traced to a $\theta \gg I$ approximation made in Eq.~(3) of \citet{ward2004II}.
Since $\theta_2 \approx I$ for $\eta \gg 1$ (Fig.~\ref{fig:cs_locs}), this
approximation is not always valid.}.

For $I = 5^\circ$, $\eta \sim \eta_{\rm c}$ Eq.~\eqref{eq:ad_constr} gives
$\epsilon \lesssim 0.085$. Since our criterion is only a weak condition for
diabaticity, we use $\epsilon = 3 \times 10^{-4}$ in our ``adiabatic''
calculations below. We explore the consequences of nonadiabatic evolution in
Section~\ref{s:nonad}.

@@

We consider the evolution of a system with arbitrary initial spin-disk
misalignment angle $\theta_{\rm sd, i}$ and initial $\eta_{\rm i} \gg 1$. We are
interested in the final spin obliquities $\theta_{\rm f}$ after $\eta$ gradually
decreases to $\eta_{\rm f} \ll 1$ (i.e.\ after the disk has dissipated to a
negligible mass). Note that when $\eta_{\rm i} \gg 1$, $\uv{l}$ precesses
around $\uv{l}_{\rm d}$ much faster than the spin-orbit precession
($\abs{\omega_{\rm ld}} \gg \abs{\omega_{\rm sl}}$), and the spin obliquity
$\theta$ varies rapidly. It is thus more appropriate to use $\theta_{\rm sd, i}$
rather than $\theta$ to specify the initial spin orientation. We explore the
entire range $\theta_{\rm sd, i} \in [0, \pi]$ and choose $\epsilon = 3 \times
10^{-4}$ so that the system evolves almost adiabatically (see above).

@@

In Section~\ref{s:ad}, we have examined the spin axis evolution in the limit
where $\epsilon \ll 1$ [see Eq.~\eqref{eq:ad_constr}] and the evolution is
mostly adiabatic (except at separatrix crossings). We now consider nonadiabatic
effects.

\subsection{Transition to Non-adiabaticity: Results for $\epsilon\lesssim
1$}\label{ss:transition}

@@

\subsection{Non-adiabatic Evolution: Result for $\epsilon\gtrsim 1$}

In general, numerical calculations are needed to determine the non-adiabatic
obliquity evolution ($\epsilon\gtrsim 1$). However, some analytical results can
be still be obtained when the obliquity remains small throughout the evolution.

We start from Eq.~\eqref{eq:dsdt_base}, which governs the evolution of the spin
axis. We choose coordinate axes such that $\uv{l} = \uv{z}$ and $ \uv{l}_{\rm d}
= \uv{z} \cos I + \uv{x}\sin I$, giving
\begin{equation}
    \rd{\uv{s}}{\tau} = \s{
        \p{\eta \cos I - \cos\theta}\uv{z} + \eta \sin I \,\uv{x}} \times \uv{s}.
\end{equation}
%Now, let's assume that $s_{\rm z} \approx \cos I$ throughout the evolution of
%$\uv{s}$ (note that $\theta_{\rm sd, i} = 0$ implies the initial $s_{\rm z} = \cos I$).
Define $S = {\hat s}_{\rm x} + i{\hat s}_{\rm y}$, we find
\begin{equation}
    \rd{S}{\tau} = i\p{\eta\cos I - \cos\theta}S - i \eta \sin I\cos\theta.\label{eq:nonad_ode}
\end{equation}
To proceed, we assume $\theta$ is small and set $\cos\theta \simeq 1$. Equation
(\ref{eq:nonad_ode}) can then be solved explicitly, starting from the initial
value $S(\tau_{\rm i})$:
\begin{equation}
    S(\tau)e^{-i\Phi(\tau)} - S(\tau_{\rm i})
        \simeq -i\sin I\int_{\tau_{\rm i}}^\tau \eta(\tau')
            e^{-i\Phi(\tau')}\;\mathrm{d}\tau',
\end{equation}
where
\begin{equation}
    \Phi(\tau) \equiv \int_{\tau_{\rm i}}^\tau \p{\eta(\tau') \cos I -1}\;
        \mathrm{d}\tau'.
\end{equation}
We now invoke the stationary phase approximation, so that $\Phi(\tau)\simeq
\Phi(\tau_0) + (1/2)\ddot\Phi (\tau_0)(\tau - \tau_0)^2$, where $\tau_0$ is
determined by $\dot\Phi=0$ or $\eta_0=1/\cos I$. We then find, for $\tau\gg
\tau_0$,
\begin{equation}
    S(\tau)e^{-i\Phi(\tau)}-S(\tau_{\rm i})\simeq -i\sin I\,\eta(\tau_0)\, e^{-i\Phi(\tau_0)}
    \p{\frac{2\pi}{i \ddot\Phi(\tau_0)}}^{1/2}.
\end{equation}
Using $\dot\eta = -\epsilon \eta$ [Eq.~\eqref{eq:deta_dt}] and
$\ddot\Phi(\tau_0) = \dot\eta(\tau_0)\cos I = -\epsilon$, we have
\begin{equation}
    S(\tau)e^{-i\Phi(\tau)}-S(\tau_{\rm i})
        \simeq -i^{3/2}\tan I\, e^{-i\Phi(\tau_0)}
            \p{\frac{2\pi}{\epsilon}}^{1/2}.
\end{equation}
If the initial obliquity is negligible ($\theta_{\rm i} = \abs{S(\tau_{\rm
i})} \ll 1$), the final obliquity is given by
\begin{equation}
    \theta_{\rm f}\simeq \sqrt{\frac{2\pi}{\epsilon}}\tan I.\label{eq:nonad_q_f}
\end{equation}
This expression is valid only if $\theta_f$ is small (recall that we have used
the approximation $\cos\theta\simeq 1$ in the above derivation).

We can generalize the above analytic result to nonzero initial spin-disc
angle $\theta_{\rm sd, i}$ by assuming the amplitude
of libration of $\uv{s}$ about $\uv{l}_{\rm d}$ (namely $\theta_{\rm sd, i}$)
is equal to the range of final obliquities when the disk dissipates quickly.
Intuitively, this corresponds to ``freezing in'' the variations in initial
obliquity over a single libration cycle to the final obliquity. Thus we expect
\begin{equation}
  \theta_{\rm f} - \sqrt{\frac{2\pi}{\epsilon}} \tan I
  \in \s{-\theta_{\rm sd, i}, \theta_{\rm sd,i}}.\label{eq:nonad_q_f_dist}
\end{equation}

In Fig.~\ref{fig:nonad_3_ensemble}, we show the numerical result of $\theta_f$
vs $\theta_{\rm sd,i}$ for $I = 5^\circ$ and $\epsilon = 0.3$. We see that
Eq.~(\ref{eq:nonad_q_f_dist}) provides good lower and upper bounds of the final
obliquity for $\theta_{\rm sd,i} \lesssim 20^\circ$.

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{plots_diskdisp/3_ensemble_05_05.png}
    \caption{$\theta_{\rm  f}$ vs $\theta_{\rm sd, i}$ for $I=5^\circ$ and
    $\epsilon = 0.3$ (firmly in the nonadiabatic regime). The blue dots
    represent numerical result, and the two red lines show the analytical lower
    and upper limits given by
    Eq.~\eqref{eq:nonad_q_f_dist}.}\label{fig:nonad_3_ensemble}
\end{figure}

Figures~\ref{fig:nonad_3_scan} and~\ref{fig:nonad_3_scan_20} show $\theta_f$ as
function of $\epsilon$ for $\theta_{\rm sd,i} = 0$ and $I = 5$, $20^\circ$. We
see that the agreement between the numerical results and
Eq.~\eqref{eq:nonad_q_f} is excellent, particularly for the $I = 5^\circ$ case.
The agreement when $I = 20^\circ$ is poorer as the initial obliquity
$\theta_{\rm i} = I$ is larger, violating one of our key assumptions. For
$\epsilon \ll 1$, we find $\theta_f \simeq 90^\circ$, in agreement with the
result of adiabatic evolution (see Fig.~\ref{fig:ad_ensemble}).

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{plots_diskdisp/3scan.png}
    \caption{Final obliquity $\theta_{\rm  f}$ as a function of $\epsilon$ for
    $\theta_{\rm sd,i} = 0$ and $I = 5^\circ$. The shaded area, bordered by the
    black line, corresponds to the adiabatic regime estimated by
    Eq.~\eqref{eq:ad_constr}. The blue dots are numerical results, and the red
    line corresponds to Eq.~\eqref{eq:nonad_q_f}, which is in good agreement
    with numerical results for $\epsilon \gtrsim 0.1$ (the nonadiabatic regime);
    $\theta_{\rm f} \simeq  90^\circ$ in the adiabatic regime ($\epsilon\ll
    1$).}\label{fig:nonad_3_scan}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{plots_diskdisp/3scan_20.png}
    \caption{Same as Fig.~\ref{fig:nonad_3_scan} but for $I=20^\circ$.
    }\label{fig:nonad_3_scan_20}
\end{figure}

@@

To examine stability of each CS, it proves very easy to handle them generally.
We linearize about an equilibrium located at $\phi_{\rm cs} = 0, \pi$ but
arbitrary $\theta_{\rm cs}$. Linearizing Eqs.~\eqref{se:H_eom} about $\phi =
\phi_{\rm cs} + \delta \phi, \theta = \theta_{\rm cs} + \delta \theta$ gives
\begin{subequations}\label{se:H_eom_lin}
    \begin{align}
        \rd{\delta \phi}{t} &= \sin \theta_{\rm cs} \delta \theta
            \mp \eta \frac{\sin I}{\sin^2\theta_{\rm cs}} \delta \theta,\\
        \rd{\delta \theta}{t} &= \pm \eta \sin I \delta \phi.
    \end{align}
\end{subequations}
Note that the positive sign corresponds to $\phi_{\rm cs} = 0$. Eliminating
$\delta \theta$ gives
\begin{align}
    \rtd{\delta \phi}{t} &= \p{\sin \theta_{\rm cs}
        \mp \eta \sin I\csc^2\theta}\p{\pm \eta \sin I} \delta
            \phi,\\
        &\equiv \lambda^2\delta \phi.\label{eq:lambda2}
\end{align}
A plot of $\lambda^2$ for each of the CSs is given in Fig.~\ref{fig:lambda2}.
From the plot, it is clear that only CS4 is a saddle point, while the other
three are centers (stable). The local libration frequency for these points is
just
\begin{align}
    \omega_{\rm lib} &= \sqrt{-\lambda^2},\\
        &= \sqrt{\p{\sin \theta_{\rm cs}
            \mp \eta \sin I \csc^2\theta}\p{\mp \eta \sin I}}.
\end{align}
The libration period is related simply $T_{\rm lib} = \frac{2\pi}{\omega_{\rm
lib}}$.
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{plots_diskdisp/2_lambdas.png}
    \caption{$\lambda^2$, given by Eq.~\eqref{eq:lambda2}, evaluated at each of
    the Cassini States. The vertical axis is rescaled for clarity. Note that CS4
    is a saddle point ($\lambda^2 > 0$) when it exists while all others are
    stable ($\lambda^2 < 0$). The horizontal dashed line is the instability
    boundary $\lambda^2 = 0$ while the vertical dashed line labels $\eta =
    \eta_{\rm c}$.}\label{fig:lambda2}
\end{figure}

>>>>

\begin{figure}
    \includegraphics[width=0.5\textwidth]{plots_diskdisp/2_cs_locs.png}
    \caption{Cassini state obliquities as a function of $\eta$ for $I =
    5^\circ$. The thin vertical dashed line indicates $\eta_{\rm c}$ ($= 0.766$
    for $I = 5^\circ$), where CS1 and CS4 merge and annihilate, and the thin
    horizontal dashed lines indicate $\theta = I$ and $I - 180^\circ$, the
    asymptotic values for CSs 2 and 3 for $\eta \gg \eta_{\rm
    c}$.}\label{fig:cs_locs}
\end{figure}

@@

As shown in Appendix~\ref{ss:eigens}, libration about CS2 is slower than that
about CS1 or CS3. As such, it has the smallest characteristic frequency in the
system. The weak adiabaticity criterion can be recast as requiring that $\eta$
vary more slowly than this libration frequency \citep[see][]{ward2004II}, i.e.\
$t_{\rm d}^{-1} \lesssim \omega_{\rm lib} / 2\pi$. In terms of the dimensionless
parameter $\epsilon$
[Eq.~\eqref{eq:eps_def}], this adiabaticity condition is
\begin{equation}
    \epsilon \lesssim \epsilon_{\rm c} \equiv
        \frac{1}{2\pi}\sqrt{\eta\sin I \sin \theta_2
            \p{1 + \eta \sin I \csc^3 \theta_2}},
            \label{eq:ad_constr}
\end{equation}
where $\theta_2$ is the obliquity at CS2. This formula differs from that given
in \citet{millholland_disk}, where the $\csc^3\!\theta_2$ term is neglected and
the square root is missing\footnote{The missing $\csc^3 \theta_2$ term can be
traced to a $\theta \gg I$ approximation made in Eq.~(3) of \citet{ward2004II}.
Since $\theta_2 \approx I$ for $\eta \gg 1$ (Fig.~\ref{fig:cs_locs}), this
approximation is not always valid.}.

For $I = 5^\circ$, $\eta \sim \eta_{\rm c}$ Eq.~\eqref{eq:ad_constr} gives
$\epsilon_{\rm c} = 0.085$. Since our criterion is only a weak condition for
diabaticity, we use $\epsilon = 3 \times 10^{-4}$ in our ``adiabatic''
calculations below. We explore the consequences of nonadiabatic evolution in
Section~\ref{s:nonad}.

@@

We consider the evolution of a system with arbitrary initial spin-disk
misalignment angle $\theta_{\rm sd, i}$ and initial $\eta_{\rm i} \gg 1$. We are
interested in the final spin obliquities $\theta_{\rm f}$ after $\eta$ gradually
decreases to $\eta_{\rm f} \ll 1$ (i.e.\ after the disk has dissipated to a
negligible mass). Note that when $\eta_{\rm i} \gg 1$, $\uv{l}$ precesses
around $\uv{l}_{\rm d}$ much faster than the spin-orbit precession
($\abs{\omega_{\rm ld}} \gg \abs{\omega_{\rm sl}}$), and the spin obliquity
$\theta$ varies rapidly. It is thus more appropriate to use $\theta_{\rm sd, i}$
rather than $\theta$ to specify the initial spin orientation. We explore the
entire range $\theta_{\rm sd, i} \in [0, \pi]$ and choose $\epsilon = 3 \times
10^{-4}$ (see above).

@@

In Section~\ref{s:ad}, we have examined the spin axis evolution in the limit
where $\epsilon \ll \epsilon_{\rm c}$ [see Eq.~\eqref{eq:ad_constr}] and the
evolution is mostly adiabatic (except at separatrix crossings). We now consider
nonadiabatic effects.

\subsection{Transition to Non-adiabaticity: Results for $\epsilon\lesssim
\epsilon_{\rm c}$}\label{ss:transition}

@@

\subsection{Non-adiabatic Evolution:
Result for $\epsilon\gtrsim \epsilon_{\rm c}$}

In general, numerical calculations are needed to determine the non-adiabatic
obliquity evolution ($\epsilon\gtrsim \epsilon_{\rm c}$). However, some
analytical results can be still be obtained when the obliquity change is small.

We start from Eq.~\eqref{eq:dsdt_base}, which governs the evolution of the spin
axis. We choose coordinate axes such that $\uv{l} = \uv{z}$ and $ \uv{l}_{\rm d}
= \uv{z} \cos I + \uv{x}\sin I$, giving
\begin{equation}
    \rd{\uv{s}}{\tau} = \s{
        \p{\eta \cos I - 1}\uv{z} + \eta \sin I \,\uv{x}} \times \uv{s}.
\end{equation}
Define $S = {\hat s}_{\rm x} + i{\hat s}_{\rm y}$, we find
\begin{equation}
    \rd{S}{\tau} = i\p{\eta\cos I - 1}S
        - i \eta \sin I\cos\theta.\label{eq:nonad_ode}
\end{equation}
To proceed, we assume the obliquity is roughly constant, $\cos \theta \approx
\cos \theta_{\rm i}$. Eq.~\eqref{eq:nonad_ode} can then be solved explicitly,
starting from the initial value $S(\tau_{\rm i})$:
\begin{equation}
    S(\tau)e^{-i\Phi(\tau)} - S(\tau_{\rm i})
        \simeq -i\sin I \cos \theta_{\rm i}
            \int_{\tau_{\rm i}}^\tau \eta(\tau')
            e^{-i\Phi(\tau')}\;\mathrm{d}\tau',
\end{equation}
where
\begin{equation}
    \Phi(\tau) \equiv \int_{\tau_{\rm i}}^\tau \p{\eta(\tau') \cos I
        - 1}\; \mathrm{d}\tau'.
\end{equation}
We now invoke the stationary phase approximation, so that $\Phi(\tau)\simeq
\Phi(\tau_0) + (1/2)\ddot\Phi (\tau_0)(\tau - \tau_0)^2$ where $\tau_0$ is
determined by $\dot\Phi = 0$, occuring when $\eta_0 = 1/\cos I$. We then find,
for $\tau \gg \tau_0$,
\begin{equation}
    S(\tau)e^{-i\Phi(\tau)} - S(\tau_{\rm i})
        \simeq -i \eta(\tau_0)\sin I \cos \theta_{\rm i} e^{-i\Phi(\tau_0)}
            \p{\frac{2\pi}{i \ddot\Phi(\tau_0)}}^{1/2}.
\end{equation}
Using $\dot\eta = -\epsilon \eta$ [Eq.~\eqref{eq:deta_dt}] and
$\ddot\Phi(\tau_0) = \dot\eta(\tau_0)\cos I = -\epsilon$, we have
\begin{equation}
    S(\tau)e^{-i\Phi(\tau)} - S(\tau_{\rm i})
        \simeq -i^{3/2}\tan I \cos \theta_{\rm i} e^{-i\Phi(\tau_0)}
            \p{\frac{2\pi}{\epsilon}}^{1/2}.
\end{equation}
If the initial obliquity is much smaller than the final obliquity ($\theta_{\rm
i} = \abs{S(\tau_{\rm i})} \ll \abs{S(\tau_{\rm f})}$), then we obtain
\begin{equation}
    \theta_{\rm f}\simeq \sqrt{\frac{2\pi}{\epsilon}}\tan I\cos \theta_{\rm i}.
        \label{eq:nonad_q_f}
\end{equation}
This expression is valid only if $\cos \theta \approx \cos \theta_{\rm i}$
throughout the evolution. This corresponds to the limit where $\theta_{\rm f}$
is not too much larger than $\theta_{\rm i}$, which requires $\epsilon$ not to
be too small. Numerically, we verify this is consistent with being in the
nonadiabatic regime $\epsilon \gtrsim \epsilon_{\rm c}$ (see
Figs.~\ref{fig:nonad_3_scan} and~\ref{fig:nonad_3_scan_20}).

We must next understand the above results in terms of the initial spin-disk
misalignment angle $\theta_{\rm sd, i}$. To do so, we first compute the final
obliquity when $\theta_{\rm sd, i} = 0$, corresponding to initial obliquity
$\theta_{\rm i} = \theta_2 \approx I$. The set of possible initial conditions
for nonzero $\theta_{\rm sd, i}$ is a ring of angular width $\theta_{\rm sd, i}$
about CS2. We find this ring evolves to a new ring of angular width $\theta_{\rm
sd, i}$ about the $\theta_{\rm sd, i} = 0$ prediction, so the range of final
obliquities is the $\theta$ range spanned by this final ring\footnote{The
interpretation for this procedure is that the process generating the final
obliquity predicted by Eq.~\eqref{eq:nonad_q_f} acts only on the \emph{center}
of libration of the trajectory when $\eta \gg 1$ (CS2). This is not apparent
from our derivation but reproduces the results of the numerical simulation.}.
This procedure yields prediction
\begin{equation}
    \abs{\p{\sqrt{\frac{2\pi}{\epsilon}}\tan I\cos I - \theta_{\rm sd, i}}}
    \leq \theta_{\rm f}
    \leq
    \sqrt{\frac{2\pi}{\epsilon}}\tan I\cos I + \theta_{\rm sd, i}
    .\label{eq:nonad_q_f_dist}
\end{equation}
The absolute value in the lower bound is a result of the ring of final
conditions crossing the North pole.

In Fig.~\ref{fig:nonad_3_ensemble}, we show the numerical result of $\theta_f$
vs $\theta_{\rm sd,i}$ for $I = 5^\circ$ and $\epsilon = 0.3$. We see that
Eq.~\eqref{eq:nonad_q_f_dist} provides good lower and upper bounds of the final
obliquity for $\theta_{\rm sd,i} \lesssim 45^\circ$.

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{plots_diskdisp/3_ensemble_05_05.png}
    \caption{$\theta_{\rm  f}$ vs $\theta_{\rm sd, i}$ for $I=5^\circ$ and
    $\epsilon = 0.3$ (firmly in the nonadiabatic regime). The blue dots
    represent numerical results, and the two red lines show the analytical lower
    and upper limits given by
    Eq.~\eqref{eq:nonad_q_f_dist}.}\label{fig:nonad_3_ensemble}
\end{figure}

Figures~\ref{fig:nonad_3_scan} and~\ref{fig:nonad_3_scan_20} show $\theta_f$ as
function of $\epsilon$ for $\theta_{\rm sd,i} = 0$ and $I = 5$, $20^\circ$. We
see that the agreement between the numerical results and
Eq.~\eqref{eq:nonad_q_f} is excellent, particularly for the $I = 5^\circ$ case.
The agreement when $I = 20^\circ$ is slightly poorer as $\theta_{\rm i}$ is
larger and hence less negligible in Eq.~\eqref{eq:nonad_q_f}. For $\epsilon \ll
\epsilon_{\rm c}$, we find $\theta_f \simeq 90^\circ$, in agreement with the
result of adiabatic evolution (see Fig.~\ref{fig:ad_ensemble}).

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{plots_diskdisp/3scan.png}
    \caption{Final obliquity $\theta_{\rm  f}$ as a function of $\epsilon$ for
    $\theta_{\rm sd,i} = 0$ and $I = 5^\circ$. The shaded area, bordered by the
    black line, corresponds to the adiabatic regime estimated by
    Eq.~\eqref{eq:ad_constr}. The blue dots are numerical results, and the red
    dashed line corresponds to Eq.~\eqref{eq:nonad_q_f}, which is in good
    agreement with numerical results for $\epsilon > \epsilon_{\rm c} \approx
    0.1$ (the nonadiabatic regime); $\theta_{\rm f} \simeq  90^\circ$ in the
    adiabatic regime ($\epsilon\ll \epsilon_{\rm c}$).}\label{fig:nonad_3_scan}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{plots_diskdisp/3scan_20.png}
    \caption{Same as Fig.~\ref{fig:nonad_3_scan} but for $I=20^\circ$. The right
    shaded area is where $\theta_{\rm f}$ predicted by Eq.~\eqref{eq:nonad_q_f}
    violates the assumption $\theta_{\rm f} \gg \theta_{\rm
    i}$.}\label{fig:nonad_3_scan_20}
\end{figure}

@@

To examine stability of each CS, it proves very easy to handle them generally.
We linearize about an equilibrium located at $\phi_{\rm cs} = 0, \pi$ but
arbitrary $\theta_{\rm cs}$. Linearizing Eqs.~\eqref{se:H_eom} about $\phi =
\phi_{\rm cs} + \delta \phi, \theta = \theta_{\rm cs} + \delta \theta$ gives
\begin{subequations}\label{se:H_eom_lin}
    \begin{align}
        \rd{\delta \phi}{t} &= \sin \theta_{\rm cs} \delta \theta
            \mp \eta \frac{\sin I}{\sin^2\theta_{\rm cs}} \delta \theta,\\
        \rd{\delta \theta}{t} &= \pm \eta \sin I \delta \phi.
    \end{align}
\end{subequations}
Note that the positive sign corresponds to $\phi_{\rm cs} = 0$. Eliminating
$\delta \theta$ gives
\begin{align}
    \rtd{\delta \phi}{t} &= \p{\sin \theta_{\rm cs}
        \mp \eta \sin I\csc^2\theta}\p{\pm \eta \sin I} \delta
            \phi,\\
        &\equiv \lambda^2\delta \phi.\label{eq:lambda2}
\end{align}
A plot of $\lambda^2$ for each of the CSs is given in Fig.~\ref{fig:lambda2}.
From the plot, it is clear that only CS4 is a saddle point, while the other
three are centers (stable). The local libration frequency for these points is
just
\begin{align}
    \omega_{\rm lib} &= \sqrt{-\lambda^2},\\
        &= \sqrt{\p{\sin \theta_{\rm cs}
            \mp \eta \sin I \csc^2\theta}\p{\mp \eta \sin I}}.
\end{align}
The libration period is related simply $T_{\rm lib} = \frac{2\pi}{\omega_{\rm
lib}}$.
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{plots_diskdisp/2_lambdas.png}
    \caption{$\lambda^2$, given by Eq.~\eqref{eq:lambda2}, evaluated at each of
    the Cassini States. The vertical axis is rescaled for clarity. Note that CS4
    is a saddle point ($\lambda^2 > 0$) when it exists while all others are
    stable ($\lambda^2 < 0$). The thin horizontal dashed line is the instability
    boundary $\lambda^2 = 0$ while the thin vertical dashed line labels $\eta =
    \eta_{\rm c}$.}\label{fig:lambda2}
\end{figure}
